---
title: "2x2 repeated measures ANOVA with RS3 data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following markdown analyses data from the RS3 in R. The results (main effects, interactions and follow up t-tests) are the same as JASP and SPSS.

```{r load packages, results = 'hide', message = FALSE}
# rm(list = ls()) # clears environment

library(ggplot2) # plotting interaction plots
library(Rmisc) # for confidence intervals on the interaction plots
library(tidyr) # manipulating dataframes
library(dplyr) # manipulating dataframes
library(ez) # allows me to compute repeated ANOVA that matches JASP output
library(car) # allows me to use Anova() function to get summary of main effects and interactions
library(sjstats) # computes partial eta squared effect size
library(effsize) # computes cohen's D for follow up t-tests
library(apaTables) # for apa tables
```

Here I am loading the packages I'll need for the analysis. I have put a note about why each one is needed. There are many ways to compute ANOVAs in R and there are many different packages you can use to do so. However I have found that the "ez" packages works best for a repeated measures design. I explain in the following markdown. 

```{r loading in data as csv file}
# work computer working directory
setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/teachingcode")

# load datset
temp = list.files(pattern = c("RS3_dataset", "*.csv")) # list all CSV files in the directory
myfiles = lapply(temp, read.csv) # read these CSV in the directory
workingdata <- do.call(rbind.data.frame, myfiles) # convert and combine the CSV files into dataframe
```

Here I am grabbing the data from my working directory

```{r descriptive statistics}
# Here I'm generating a participant ID, this will be important later
workingdata <- workingdata %>%
  mutate(ppid = row_number())

# computing histograms and normality curves for each condition
ggplot(workingdata %>%
         gather("NoAS_Ver", "NoAS_VerDemo", "AS_Ver", "AS_VerDemo", key = "condition", value = "correctpairs"), aes(x = correctpairs)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(
		fun = dnorm, 
		args = with(workingdata %>%
		              gather("NoAS_Ver", "NoAS_VerDemo", "AS_Ver", "AS_VerDemo", key = "condition", value = "correctpairs"), c(mean = mean(correctpairs), sd = sd(correctpairs)))) + 
  facet_wrap( ~ condition)

# descriptives for age
range(workingdata$Age)
mean(workingdata$Age)
sd(workingdata$Age)
```

Firstly I'm creating a new variable "ppid" to indicate the participant number. This will be important when the ANOVA in R attempts to calculate the degrees of freedom. 

Histograms are plotted with normal distributions to show how the data is distributed. In this case, things seem fairly normally distributed.

```{r manipulating dataframe}
# spliting conditions into factors and their levels
split_data <- workingdata %>%
  gather("NoAS_Ver", "NoAS_VerDemo", "AS_Ver", "AS_VerDemo", key = "condition", value = "correctpairs") %>%
  separate(condition, into = c("task", "presentation"))

head(split_data, n = 5)
tail(split_data, n = 5)

# sets our factors as a factor data types
split_data$task <- as.factor(split_data$task)
split_data$presentation <- as.factor(split_data$presentation)
split_data$ppid <- as.factor(split_data$ppid)
```

In JASP, each column represents a condition. In R, I need the columns to refer to the different factors, with each row being an observation. I use the gather function to create a condition and value column. I then use the separate function to seperate the conditions into "task" and "presentation". The head and tail function demonstrate what this data frame looks like now.

I then ensure the factors in this design are set to factor data types.

```{r running the ANOVA the WRONG way}
# running the ANOVA this way generates the wrong f values and sums of squares
result_wrong <- aov(correctpairs ~ task * presentation, data = split_data)
Anova(result_wrong, type = 3)
```

This is how I first tried to run the ANOVA. However as you can see, the sum of squares, f values and degrees of freedom (DF) do not match JASP and SPSS output (despite using type 3 sum of squares). After many nights of searching, I realised the problem is the DFs.

```{r number of rows are be equal to the DFs}
nrow(split_data)
```

The DFs are the number of observations that are allowed to vary in the statistical calculation and are calculated as the N - 1. However because I have arranged the dataframe so that each column is a factor, the dataframe is telling the ANOVA that I have 1780 observations. I actually have 445 (the number of participants).

```{r running the ANOVA the RIGHT way}
# running the ANOVA this way generates the correc f values and sum of sqaures
result_correct <- ezANOVA(split_data, dv = correctpairs, wid = ppid, within = .(task, presentation), type = 3, detailed = TRUE, return_aov = TRUE)
result_correct
```

If I use the ezANOVA function, I can specify the number of cases I actually have with the "wid" arugument. I assign this the ppid variable made earlier. This means that the DFs can be correctly calculated, and now the sum of squares (SSn), F value and DFs match the JASP and SPSS output.

Importantly, I input "return_aov = TRUE". This is important for the effect sizes.

```{r effect size}
# effect sizes for the correct ANOVA method
eta_sq(result_correct$aov, partial = TRUE)
```

The effect size function only takes "aov" objects. Because I returned an "aov" object from the ezANOVA, I can use this to calculate the effect size. 

```{r estimated marginal means for the main effects}
model.tables(result_correct$aov, "mean", se = TRUE)
```

I then also use the "aov" object from the ezANOVA to calculate the estimated marginal means for the main effects. Again, these match the ones from the SPSS output.

```{r interaction plot}
# computes standard error and 95% confidence intervals
CIerror_bar <- summarySE(split_data, measurevar = "correctpairs", groupvars = c("task", "presentation"))

ggplot(data = CIerror_bar, aes(x = presentation, color = task, group = task, y = correctpairs)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = correctpairs - se, ymax = correctpairs + se), width = .1)
```

Here I have an interaction plot that visualises the interaction. I have also plotted 95% confidence intervals around each mean.

```{r follow up paired sample t tests}
workingdata <- split_data %>%
  unite("condition", task, presentation, sep = "_") %>%
  spread(condition, correctpairs)
```

To compute the t-tests, I re-order the dataframe so that each condition is a column.

```{r t-tests and effect sizes}
# t-test and effect size for AS and no AS in the verbal condition
t.test(workingdata$NoAS_Ver, workingdata$AS_Ver, paired = TRUE)
cohen.d(workingdata$NoAS_Ver, workingdata$AS_Ver, paired = TRUE)

# t-test and effect size for AS and no AS in the verbal condition
t.test(workingdata$NoAS_VerDemo, workingdata$AS_VerDemo, paired = TRUE)
cohen.d(workingdata$NoAS_VerDemo, workingdata$AS_VerDemo, paired = TRUE)
```

I can then compute t-tests and cohen's D effect sizes for AS and no AS in the verbal and and verbal + demo presentation factor levels.

```{r means and anova with table output in word document}
setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/teachingcode")

apa.2way.table(presentation, task, correctpairs, data = split_data, filename = "anova_means_table.doc", table.number = 1,
  show.conf.interval = TRUE, show.marginal.means = TRUE)

apa.ezANOVA.table(result_correct, table.title = "Within subjects effects", table.number = 2, filename = "anova_table.doc")
```

Finally, a nice little touch. The apaTables package allows me to save the mean and SDs from the conditions and the ANOVA tables as an APA formatted table in a word document. The package also includes functions that do this for over tests such as regressions and one-way ANOVAs.
